%--------------------------------------------------------------------------------------------
% Document revision history
%--------------------------------------------------------------------------------------------

\chapter{Revision History}


\section*{11 May 2019}

\begin{itemize}
\item Update the default value for \namelist{mnl:config_o3climatology} to match the v6.3 release: the default is now `true'.
\end{itemize}

\section*{17 April 2018}

\begin{itemize}
\item Minor updates for MPAS v6.0, including new \namelist{inl:config_topo_data} option and new \field{initial_time} field.
\end{itemize}

\section*{22 March 2018}

\begin{itemize}
\item Update Appendix \ref{chap:atm_fields} with correct units for \field{scalars_tend}.
\end{itemize}

\section*{1 August 2017}

\begin{itemize}
\item Update documentation in Chapter \ref{chap:mpas_build_instructions} for building with PIO 2.x versions
\item Correct other minor typographical errors and unclear wording
\end{itemize}

\section*{12 May 2017}

\begin{itemize}
\item Add documentation for new \namelist{inl:config_extrap_airtemp} initialization namelist option
\item Remove documentation for config\_smdiv\_p\_forward model option, which was removed in v5.1
\item Add note that \namelist{mnl:config_len_disp} is also used by 3-d divergence damping in v5.1
\item Change default value of \namelist{mnl:config_smdiv} model option to 0.1, reflecting changes in v5.1
\item Other minor changes to match the MPAS-Atmosphere v5.1 release
\end{itemize}

\section*{23 December 2016}

\begin{itemize}
\item Added description of the new `convection\_permitting' physics suite, as well as
a list of all possible physics parameterizations, to Chapter 6.
\item Updated the chapter on building MPAS to mention specific versions of PIO that are known
to work with MPAS, information on the method used to build single-precision executables in v5.0,
and a mention of the {\em experimental} OpenMP capability in v5.0.
\item Appendices A, B, and D are now automatically generated based on XML Registry files.
\item Many other small changes to reflect the state of the v5.0 release.
\item Minor wording changes and corrections of typographical errors throughout document.
\end{itemize}

\section*{19 May 2015}

\begin{itemize}
\item Added chapter describing the MPAS-Atmosphere physics suites introduced in MPAS v4.0.
\item Added a section on building the model with single-precision reals.
\item Updated the I/O chapter to include the new io\_type option introduced in MPAS v4.0.
\item Updated a few namelist options to match MPAS v4.0 code.
\item Minor wording changes and corrections of typographical errors throughout document.
\end{itemize}

\section*{18 November 2014}

\begin{itemize}
\item Added chapter describing the MPAS runtime I/O system available in MPAS v3.0.
\item Updated the chapter on running MPAS-Atmosphere to match MPAS v3.0 code.
\item Updated a few namelist options to match MPAS v3.0 code.
\item Minor wording changes throughout document.
\end{itemize}

\section*{13 June 2013}

Initial version.
